---
title: "simsalRbim - Worth Evaluation"
author: 
  - Steven R. Talbot
  - Dana Pfefferle
  - Lars Lewejohann
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simsalRbim - Worth Evaluation}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height = 5.5
)
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(simsalRbim)
library(knitr)
library(dplyr)
library(kableExtra)
```



```{r include=FALSE}
dat        <- ZickeZacke

simOpt     <- "HoiHoiHoi"
GT         <- c( "Huehner", "Kacke",  "Zicke","Zacke" )

predat     <- bimpre (dat=dat, GT=GT, simOpt=simOpt, deviation=0, minQuantity=0)

worth      <- bimworth(ydata    = predat,
                       GT       = GT,
                       simOpt   = simOpt,
                       randOP   = FALSE,
                       intrans  = FALSE,
                       showPlot = FALSE)
```



## Worth Evaluation
The `bimeval` function compares the preferences of each subject and calculates
the mean consensus error. Because this can be a bit abstract, we provide an 
example below.

### Example item: 'HoiHoiHoi'
In our dataset 4 individuals have ranked all items except "HoiHoiHoi". Moreover,
those that have rated "HoiHoiHoi" were not in total agreement whether or not it 
is preferred in binary comparisons to other items. The amount of disagreement is
termed consensus error, with a consensus error of 0% meaning all individuals
agree on a rank position and a consensus error of 100% meaning all individuals
disagree. Let’s say, we want to assess the consensus error of the newly entered
‘HoiHoiHoi’ item in the ZickeZacke dataset. For this, we consider all item 
pairings that include ‘HoiHoiHoi’.

The table below provides the quantities (e.g., amount drunk) for items A and B
if available, and the result of the preference in relation to quantity A, i.e., 
1 (preferred item A), 0 (tie), -1 (prefered item B). Item pairings that are not
provided in the data are filled in as ties (note, that the quantities are zero 
in these cases, while those ties that already occurred in the original data 
still show quantities).

```{r echo=FALSE}
predat_mod     <- predat#[predat$sim==FALSE, ]

predat_mod[predat_mod$optionB == "HoiHoiHoi", c("subjectID","optionA","optionB",
                                        "quantityA","quantityB","result")] %>%
  kbl() %>%
  kable_styling()
```

For example, in the comparison ‘Zacke’ vs ‘HoiHoiHoi’ the table above shows that
from four subjects only one (‘vier’) preferred the item ‘Zacke’ over the item 
‘HoiHoiHoi’, result = 1. The other subjects prefered ‘HoiHoiHoi’ over ‘Zacke’,
result = -1. This is converted into percentages, so that 25% of the total 
subjects preferred option A (‘Zacke’) and 75% option B (‘HoiHoiHoi’).

This calculation is applied to all item pairings in the dataset, resulting in 
the following table. With the ‘HoiHoiHoi’ item all item combinations are tested,
resulting in 4 test combinations. Further, missing subjects are filled as ties.
Since subjectID = 'drei' is missing for the test combinations 'Zacke' vs 
'HoiHoiHoi' and 'Zicke' vs 'HoiHoiHoi', two ties are introduced. In total, this
results in a 4 subjects x 4 item combinations table.


```{r echo=FALSE}
optionlist <- c(simOpt, GT)

# compute the decision matrix for all data combinations in the predat object
# no side is present
TT <- NULL
for(i in 1:length(unique(predat$test))){
  tt <- data.frame(test           = as.character(unique(predat[predat$test==unique(predat$test)[i], "test"])),
                   pct_first      = ifelse(predat[predat$test==unique(predat$test)[i],    ]$result== 0,
                                           0,
                                           sum(predat[predat$test==unique(predat$test)[i],    ]$result== 1) /
                                             length(unique(predat$subjectID))*100),
                   pct_second     = ifelse(predat[predat$test==unique(predat$test)[i], ]$result ==0,
                                           0,
                                           100 - (sum(predat[predat$test==unique(predat$test)[i], ]$result== 1) /
                                                    length(unique(predat$subjectID))*100)))
  TT <- rbind(TT, tt)
}


AA <- TT
colnames(AA) <- c("test","option A (%)", "option B (%)")
AA <- AA[grepl('HoiHoiHoi', AA$test, fixed=TRUE),]
rownames(AA) <- NULL

AA %>%
  kbl() %>%
  kable_styling()

```

In the example ‘Kacke’ vs ‘HoiHoiHoi’, 100% of the subjects preferred option A 
(‘Huehner’), etc.


### Calculation of the Consensus Error
The ‘consensus error’ (CE) indicates the percent of (dis)agreement in preference
within a given item pairing.

We set the threshold for deciding whether option A or B is preferred to 50% 
(default setting) and calculate the deviation of the smallest percentage from 
it (deviation = delta). (Note: the threshold can be changed in the `bimpre` 
functions using the function objects `deviation` and/or `minQuantity`). 

The following pseudocode illustrates how the deviation delta was calculated.

```{r eval=FALSE, echo=TRUE}
# Pseudococde - deviation calculation
for(i in 1:item_combinations){ # e.g., ZackeHoiHoiHoi
  if( item[i, "option A" ] > item[i, "option B" ] ){
    delta  <- (50 - item[i, "option B" ])
  }else{
    delta  <- (50 - item[i, "option A" ])
  }
}
```

The delta is averaged, standardized, and subtracted from 50%, resulting
in the consensus error (CE in %):

\begin{align*}
CE = 50-(\sum_{i=1}^n delta_i/n)/50*100\\
\end{align*}

The mean delta of all 'HoiHoiHoi' items is 9.38%:

((50+50+50+50) + (50+50+50+50)+ (50+50+50+50) + (25+25+25+50))/16 = 725/16 = 
45.31% deviation.

Then we calculate 50%-45.31%=4.69% deviation. This is standardized to 50,
resulting in a CE of 4.69%/50*100=9.38% which is also shown in the following
plot.

Therefore, the item ‘HoiHoiHoi’ shows a consensus error of 9.38% indicating that
9.38% of test subjects disagreed on the positioning of the item. This is also 
displayed in the corresponding plot of the `bimeval` function.


```{r echo=FALSE}
options(warn=-1)
w_errors   <- bimeval(ydata     = predat, 
                      worth     = worth, 
                      GT        = GT,
                      simOpt    = simOpt,
                      filtersim = FALSE, 
                      ylim      = c(0,0.8),
                      showPlot  = TRUE )
options(warn= 0)
```



