---
title: "simsalRbim Examples"
author: "Steven R. Talbot"
output: 
    rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simsalRbim Examples}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height = 5
)
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(simsalRbim)
library(knitr)
library(kableExtra)
```

## Preference Tests
For this Vignette, we will use the artificial ZickeZacke data. 
This is a 'perfect' set of preference tests that comes with this package, in
which every subject was tested with every provided item. However, the data
contain one item that was not tested with all subjects and also contains two
ties ('HoiHoiHoi'). Let's see if we can evaluate the item's position. For this 
purpose, we will employ three different strategies which are covered by the 
following examples.


### Example 1 - The calculation of worth values
First, the data are loaded. In this example, internal data are used. You can
use the `bimload` function to import data from a variety of formats.

Next, an item that shall be tested is required (simOpt), and a Ground Truth (GT)
of items that it is tested with. These entries depend on the data structure. Any
items that are not listed here, will not be used in the analysis. This may lead
to incomplete conclusions. So be careful what you put in here...

This information is then used in the `bimpre` function which does some 
pre-processing. Also, ties are identified. As you can see from the example 
below, the ties were marked in the 'tie' column.


```{r}
dat        <- ZickeZacke

simOpt     <- "HoiHoiHoi"
GT         <- c("Zicke", "Zacke", "Huehner", "Kacke" )

predat     <- bimpre(dat=dat, GT=GT, simOpt=simOpt)

```

 
In the next step, we will calculate the worth values of the item. This uses a
non-linear GNM model to model the data and transforms the estimates into worth
values. In the `showPlot` object, you can choose between the worth values or the 
model coefficients ("coef").

Note, that when `randOP` is set to TRUE, the attribution of ties in terms of the
dependent results variable in the GNM model will be random. Each time this 
function is executed, the worth values will change. This is an integral part of
this package and will be used in the simulations later.

```{r}
worth      <- bimworth(ydata    = predat,
                       GT       = GT,
                       simOpt   = simOpt,
                       randOP   = FALSE,
                       showPlot = "worth")
```

As you can see, the item "HoiHoiHoi" appears to be marginally better than
"Zacke". But what can we actually say about the general data quality of this 
plot? Maybe only a fraction of individuals were tested, or just some items and
not all of them. When data are large this becomes difficult to evaluate.

Therefore, we will use the `bimeval` function to evaluate the data together with
the worth values. The results will be visualized as a bubble plot.

Also, with these data, there will be two **warnings**. This is intentional.

```{r}
w_errors   <- bimeval(ydata    = predat, 
                      worth    = worth, 
                      GT       = GT, 
                      simOpt   = simOpt,
                      showPlot = TRUE)
```

The warnings indicate that the number of subjects tested may be too low. The 
second warning shows the same for the number of items tested. To get
more reliable results, both should be increased.

The bubble plot shows the consensus error of each item. This means, the smaller
the bubble (or error), the more agreement there is between subjects regarding
the positioning of the respective item. With this tool, we can assume that the 
position of 'Zicke' at zero % error is secure. There was a full agreement of all 
subjects on the positioning of that item. However, 'HoiHoiHoi' shows 25% 
disagreement. Take this together with the warnings from above, the positioning
of this item is relatively insecure. It might be in position 2 or 3...

## Example 2 - uninformed item position simulation
From the example above we know that for the inclusion of ties, the results of 
the preference test can be randomized. In the GNM it does not matter whether the
test was binary or continuous. The preprocessing took care of that. However, for
a certain amount of combinations, this leaves some degrees of freedom 
that can be modeled in the GNM. This logic is used in the `bimUninformed` 
function to simulate an **uninformed*'** worth calculation for optimal item
positioning. 

This procedure is called 'uninformed' because it uses no *a priori* knowledge
about the transitivity of items. It is assumed that all choices in the 
preference test were correct.

For any number of randomization steps, an ANOVA with an MCP correction is used to
compare the worth values. As long as they are not significantly different, the 
items' position can be challenged. At a limited number of degrees of freedom, 
this will eventually lead to a saturation of the mean adjusted p-value. 
Starting at relatively large confidence levels, the errors will become smaller
and once the p-value hits zero, the cutoff is found. For ambiguous data, this can
mean a lot of required randomizations. **This is a heuristic!** So, be prepared
to try this out manually for your data.


```{r}
# We will run 100 randomizations in this example to find the optimal cutoff
cutoff     <- bimUninformed(ydata      = predat, 
                            GT         = GT,
                            simOpt     = simOpt, 
                            limitToRun = 100, 
                            ylim       = c(-1,2) )
cutoff$cutoff
```

The simulation shows that the optimal cutoff for the data is
at `r cutoff$cutoff` randomizations.

This result can be displayed with 95% confidence intervals using the `bimpos`
function. Here, we will insert the number of required randomizations.

```{r}
pos        <- bimpos(ydata      = predat, 
                     GT         = GT, 
                     simOpt     = simOpt, 
                     limitToRun = cutoff$cutoff, # 65
                     showPlot   = TRUE )

```

```{r echo=FALSE}
pos$simerrors %>%
  kbl() %>%
  kable_styling()
```

The table shows the estimated mean worth values with errors. All values
are sorted in the correct order. The items' positions are now clear as there are
no overlapping confidence intervals. 

From the evaluation above our confidence in the positioning of 'HoiHoiHoi' was
low. Now we can be rather sure that it is in position 2. If in doubt, increase 
the number of randomizations until the items are separated.

## Example 3 - informed item position simulation
In the last example, we will perform an **informed** simulation of the data. By 
calculating the intransitivity of the item combinations, we can estimate which
choices were wrong and include this in our evaluation. 
We will use a frequency distribution to finally decide on the tested item's
position.

```{r}
frqnc      <- bimsim(ydata       = predat,
                     GT          = GT,
                     simOpt      = simOpt, 
                     limitToRun  = 78,
                     fval        = 1, 
                     showPlot    = TRUE, 
                     ylim        = c(0,0.7))
```


```{r echo=FALSE}
frqnc$frq %>%
  kbl() %>%
  kable_styling() 
```

The table shows, that the items with a low intransitivity ratio (Iratio) are 
mostly in position 2. Therefore, for the 'HoiHoiHoi' item, we can opt for
position 2.





