---
title: "simsalRbim - Quality measures for worth evaluation"
author:
  - Steven R. Talbot
  - Dana Pfefferle
  - Ralf Brockhausen
  - Lars Lewejohann
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simsalRbim - Quality measures for worth evaluation}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height = 5.5
)
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(simsalRbim)
library(knitr)
library(dplyr)
library(kableExtra)
```



```{r include=FALSE}
dat        <- ZickeZacke

simOpt     <- "HoiHoiHoi"
GT         <- c( "Huehner", "Kacke",  "Zicke","Zacke" )

predat     <- bimpre (dat=dat, GT=GT, simOpt=simOpt, deviation=0, minQuantity=0)

worth      <- bimworth(ydata    = predat,
                       GT       = GT,
                       simOpt   = simOpt,
                       randOP   = FALSE,
                       intrans  = FALSE,
                       showPlot = FALSE)
```


### Why do we need quality measures?
When simulating scaled rankings with unknown values, it is important to 
estimate measures of fit. We provide two measures for doing this: the 
*consensus error* (CE) and the *intransitivity ratio* (Iratio). 

## Consensus Error (CE)
The ‘consensus error’ (CE) indicates the percent of (dis)agreement in preference
within a given item pairing between individuals. If all individuals agree on a
given item, the CE is zero. If individuals disagree, there is no consensus. The
range of disagreement is between one and half of the individuals who differ in
their preference for a given set of two items. The `bimeval` function compares 
the preferences of each subject and calculates the mean consensus error. 
Because this can be a bit abstract, we provide an example below.


### Example item: 'HoiHoiHoi'
In our ZickeZacke dataset, four individuals have rated all items, except 
'HoiHoiHoi', which was rated by only 3 individuals (“eins", “zwei”, “vier”). 
Moreover, those individuals that have rated 'HoiHoiHoi' were not in total
agreement whether or not 'HoiHoiHoi' was preferred in binary comparisons over
other items. The amount of disagreement is termed consensus error, with a 
consensus error of 0% meaning all individuals agree on a position, and a
consensus error of 100% means all individuals disagree. Let’s say, we want to
assess the consensus error of the newly entered ‘HoiHoiHoi’ item in the 
ZickeZacke dataset. For this, we consider all item pairings that include 
‘HoiHoiHoi’.

The table below provides the quantities (e.g., amount drunk) for items (options)
A and B if available, and the result of the preference in relation to quantity 
A, i.e., 1 (preferred item A), 0 (tie), -1 (prefered item B). Item pairings that
are not provided in the data are filled in as ties (note, that the quantities 
are zero in these cases, while those ties that already occurred in the original 
data still show quantities).

As an example, the combinations of "Kacke" with the "HoiHoiHoi" item are
highlighted. Note that subjectID "drei" has no quantities and was therefore 
added as a tie.

```{r echo=FALSE}
w_errors   <- bimeval(ydata     = predat,
                      worth     = worth,
                      GT        = GT,
                      simOpt    = simOpt,
                      showPlot  = FALSE ,
                      subtitle  = NULL,
                      ylim      = c(0,1))
AA <- w_errors$predat[, c("subjectID", "optionA", "optionB", "quantityA",
                          "quantityB", "test", "result", "tie")]
rownames(AA) <- NULL

color.me     <- which(AA$test == "KackeHoiHoiHoi")

AA %>%
  kbl() %>%
  kable_styling() %>%
  row_spec(color.me, bold = F, color = "white", background = "red")



```

For the next step in the CE calculation, the function counts how many cases of 
preference for A or B and how many ties were found. This number is multiplied
with  1/No. of subjects.  

In the example of the highlighted case above, this would be  
No. A x 1/No. of subjects = 0 x 0.25 = 0

No. B x 1/No. of subjects = 2 x 0.25 = 0.5  

To avoid distribution bias, ties are attributed equally to options A and B, 
e.g., in the data above there are 2 ties, so the calculation is:  

(No. of A x 1/No. of subjects) + ((No. ties x 1/No. of subjects)/No. of ties =
(0 x 0.25) + ((2x0.25)/2) = 0.25.

*AND*  

(No. of B x 1/No. of subjects) + ((No. ties x 1/No. of subjects)/No. of ties) =
(2 x 0.25) + ((2x0.25)/2) = 0.75.

This is repeated with all item combinations, resulting in the following table:

```{r echo=FALSE}

BB <- w_errors$preCE[, c("test","n","no_of_ties", "CE_A", "CE_B")]

BB %>%
  kbl() %>%
  kable_styling() 
```

### Calculation of the Consensus Error
The ‘consensus error’ (CE) indicates the percent of (dis)agreement in preference
within a given item pairing.

We set the threshold for deciding whether option A or B is preferred to 50% 
(default setting) and calculate the deviation of the smallest percentage from 
it (deviation = delta). (Note: the threshold can be changed in the `bimpre` 
functions using the function objects `deviation` and/or `minQuantity`). 

The following pseudocode illustrates how the deviation delta was calculated.

```{r eval=FALSE, echo=TRUE}
# Pseudococde - deviation calculation
for(i in 1:item_combinations){ # e.g., ZackeHoiHoiHoi
  if( item[i, "CE_A" ] > item[i, "CE_B" ] ){
    delta[i]  <- (50 - item[i, "CE_B" ])
  }else{
    delta[i]  <- (50 - item[i, "CE_A" ])
  }
  CE_raw      <- 50-delta
  CE          <- (50-delta)/50 * 100
}
```

With the resulting delta value, the deviation from 50% can be calculated. A 
delta value of 50 means perfect consensus. Therefore: CE_raw(A,B) = 50 - delta.

This is the *un*standardized Consensus Error.

The unstandardized error may be unintuitive. Therefore, it is standardized to 
percent range by: CE = (50-delta)/50 x 100  

This is repeated with all items, resulting in the following detailed error 
table.  

```{r echo=FALSE}

CC <- w_errors$detailed

CC %>%
  kbl() %>%
  kable_styling() 
```

The detailed error table is summarized for each item by reporting the average CE
together with the worth value. The final error table is, therefore:

```{r echo=FALSE}

DD <- w_errors$errors

DD %>%
  kbl() %>%
  kable_styling() 
```

Finally, the total CE is calculated as the average of the averaged item CEs (
here: 5 items).

**Total CE** = (25 + 31.25 + 25 + 25 + 81.25) / 5 = **37.5%**

The `bimeval` function plots this result as a bubble plot in which the CE 
determines the bubble sizes.  

```{r echo=FALSE}
w_errors   <- bimeval(ydata     = predat,
                      worth     = worth,
                      GT        = GT,
                      simOpt    = simOpt,
                      showPlot  = TRUE ,
                      subtitle  = NULL,
                      ylim      = c(0,0.65))
```


## Intransitivity
The combination of multiple bimodal choices is usually based on the assumption 
of transitivity: If item A is preferred over B, and B is 
preferred over C, then A should also be preferred when directly compared to
C. However, in real-life data, intransitive decisions occur on the individual as
well as on the group level. Therefore, the amount of intransitivity in the 
derived ranking can be used as a proxy for the quality of the ranking.

The transitivity is always checked for the whole data. The number of 
intransitive triplets is counted and standardized to the total number of
triplets. For example:

Let's say, there are 40 triplets in total and 3 triplets show intransitive
patterns. Then, the intransitivity ratio would be **Iratio = 3/40 = 0.075**.

Introducing simulated ties can lead to intransitivity or help with improving 
already present intransitive pairs. Therefore, running the simulation longer 
will increase the chances of obtaining transitive pairs for new item 
combinations, which we select for in the informed simulation. Usually, the best 
choice for a new item will be the one with the highest achieved transitivity.

This method is used in the informed simulation of the `bimsim` function. 
The results are shown in bubbleplots with the transitivity ratio as size 
parameter.

```{r echo=FALSE}
frqnc      <- bimsim(rawdat      = dat,
                     GT          = GT,
                     simOpt      = simOpt,
                     limitToRun  = 50,
                     tcut        = 0.90,
                     filter.crit ="Iratio",
                     ylim        = c(0,0.45))
```

 

